---
title: "Week 2"
author: Josh Jackson
date: "10-22-20"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = "#23395b",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)

library(tidyverse)
library(broom)
```

<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 4em 1em 4em;
}

.small .remark-code { 
  font-size: 80% !important;
}
.tiny .remark-code {
  font-size: 65% !important;
}
</style>


## Goals for today

This week we are going to circle back to much of what we had already done in week 1. The difference will be that we will be walking through an example so as to revisit some of the terms we introduced. Also, we will be doing analyses using `brms`, the primary package we will be using in class. 


