

---
## Goals for this week
Go deeper into basic linear models to become comfortable with standard regressions


---
## Prior Predictive


---
## Bayesian R2


---
## Rhat and other model components




---
## Categorical data



---
### dummy variables



---
## Factorial design









---
## Hypothesis function


---
## update function


---
## emmeans




---
## pp checks



---
## Robust regression
https://solomonkurz.netlify.app/post/robust-linear-regression-with-the-robust-student-s-t-distribution/

We are going to use a t-distribution as our likelihood. 

```{r, eval = FALSE}
b2 <- 
  brm(family = student,
      y ~ 1 + x,
      prior = c(prior(normal(0, 10), class = Intercept),
                prior(normal(0, 10), class = b),
                prior(gamma(4, 1), class = nu),
                prior(cauchy(0, 1),  class = sigma)),
      data = data)
```

---
### gamma plot


---
### fixing nu

```{r, eval = FALSE}
b4 <-
  brm(family = student,
      bf(y ~ 1 + x, nu = 4),
      prior = c(prior(normal(0, 100), class =Intercept),
                prior(normal(0, 10),  class = b),
                prior(cauchy(0, 1),   class = sigma)),
         data = o)
```




---
## Correlation preview

Introduce multivariate models




## For homework

Give model definition, ask to define likelihood, priors, etc. Then Ask how many parameters are in posterior distribution. 

Prior predictive simulations

visualizing different model effects


