---
title: "HW-2"
output: html_document
---

We are going to be working with samples a lot so this week we will be familiarizing our selves with sampling and describing samples using different packages. Because I do not want you to actually do any grid approximation, we are going to be working from a given posterior rather than calculating it. The posterior is a modified one from the example presented in week 2. Data are available from here: https://raw.githubusercontent.com/josh-jackson/bayes/master/p_grid.csv As a reminder there is a posterior for both mu and sigma

1. Modify and use the below code to sample from the posterior. Create a dataframe with 100 samples, 1000 samples and 10,000 samples for both mu and sigma
```{r,eval = FALSE}
library(tidyverse)
p_grid_samples <- 
  p_grid %>% 
  sample_n(size = XX, replace = T, weight = probability) %>% 
  pivot_longer(mu:sigma) 
```

Graph the posterior of each of these (your choice if you want to do the heatmap or the densities of each)


2. We often want to summarize our samples. For each of the three different samples we took, can you create a table of the values you think would be useful to summarize the distribution. Specifically some measure of central tendency and a measure of dispersion. Use a different combination for each set of samples. 

{tidybayes} is a package we will use a lot this semester and has some convenience functions to summarize samples. You can use your own code or utilize tidybayes functions. 
tidybayes summary stats have two parts. The first, specifies which measure of central tendency --mean median mode. 
```{r, eval = FALSE}
mode_hdi()
mean_hdi()
median_hdi()
```

The second component describes which type of summary you want. Do you want a percentile interval (aka an equal tailed interval, referred here as a quantile interval, qi) or do you want a hpdi, referred here as hdi interval. 
```{r, eval = FALSE}
mode_qi()
mode_hdi()
```


3. While we do not do p-value tests within bayes, we can approximate something akin to that by finding area certain areas under the curve equal to our p-value eg 2.5% on each side for an alpha of .05. Or 5% for a 1 tailed test. For each sample of the posterior distribution find value that demarcates the lower 8%. 


4. Using your answer to 1 and to 2 can you graph a density showing the credible interval. Only do 1 graph for mu and one for sigma. your choice on the interval. 

