---
title: "grabbag"
author: Josh Jackson
date: "12-10-20"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = "#23395b",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)

library(tidyverse)
library(broom)
library(tidybayes)
library(modelr)
```

<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 4em 1em 4em;
}

.small{ 
  font-size: 80%}
.tiny{
  font-size: 65%}
</style>

## This time

Hodgepodge:
1. Mediation & multivariate/distribution models
2. Missing data
3. IRT


---
## Simple mediation

.pull-left[
$$M  = i_M + a X + e_M$$
$$Y  = i_Y + c' X + b M + e_Y$$

]

.pull-right[
```{r, echo = FALSE, message = FALSE}

library(ggdag)

dag_coords <-
  tibble(name = c("X", "M", "Y"),
         x    = c(1, 2, 3),
         y    = c(2, 1, 2))

p1 <-
  dagify(M ~ X,
       Y ~ X + M,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "black", alpha = 1/4, size = 10) +
  geom_dag_text(color = "black") +
  geom_dag_edges(edge_color = "black") +
  scale_x_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  theme_bw() +
  theme(panel.grid = element_blank())+
  ggtitle("direct + indirect effect")

p2 <-
  dagify(Y ~ X,
       coords = dag_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "black", alpha = 1/4, size = 10) +
  geom_dag_text(color = "black") +
  geom_dag_edges(edge_color = "black") +
  scale_x_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  theme_bw() +
  theme(panel.grid = element_blank())+
  ggtitle("Total effect")

library(patchwork)

p2 | p1

```
]

---
# path analysis with lavaan


```{r}
library(lavaan)

X <- rnorm(100)
M <- 0.5*X + rnorm(100)
Y <- 0.7*M + rnorm(100)
Data <- data.frame(X = X, Y = Y, M = M)
model <- ' # direct effect
             Y ~ c*X
           # mediator
             M ~ a*X
             Y ~ b*M
           # indirect effect (a*b)
             ab := a*b
           # total effect
             total := c + (a*b)
         '
fit <- sem(model, data = Data)

```


---
.tiny[
```{r}

summary(fit)
```
]

---

```{r, message = FALSE, warning = FALSE}
library(blavaan)
  b.fit <- bsem(model, data=Data, mcmcfile = "blavaan.1")

```

---
.tiny[
```{r}
summary(b.fit)
```

]

---
# brms
```{r}
library(brms)
# describe your equations
y_model <- bf(Y ~ 1 + X + M)
m_model <- bf(M ~ 1 + X)

# simultaneously estimate
med.1 <-
  brm(family = gaussian,
      y_model + m_model + set_rescor(FALSE),
      data = Data,
      cores = 4,
      file = "med.1")

```


---
.tiny[
```{r}
summary(med.1)
```
]


---
## Indirect effects

```{r}
library(tidybayes)
get_variables(med.1)
```

```{r}
med.1 %>% 
  spread_draws(b_Y_M, b_M_X,b_Y_X) 

```


---
## calculate indirect effects
```{r}
med.1 %>% 
  spread_draws(b_Y_M, b_M_X, b_Y_X) %>% 
  mutate(indirect = b_Y_M * b_M_X) %>% 
  mutate(direct = b_Y_X) %>% 
  mutate(total = indirect + direct ) %>% 
  median_qi(indirect, direct,total)
```


---

```{r}
med.1 %>% 
  spread_draws(b_Y_M, b_M_X, b_Y_X) %>% 
  mutate(indirect = b_Y_M * b_M_X) %>% 
  mutate(direct = b_Y_X) %>% 
  mutate(total = indirect + direct ) %>% 
  select(indirect, direct, total) %>% 
  gather() %>% 
  ggplot(aes(y = key, x = value)) +
  stat_dotsinterval()
```

---
## priors for mediation
7 parameters to estimate

```{r}
get_prior(y_model + m_model + set_rescor(FALSE),
          family = gaussian,
          data = Data)
```


---
```{r}
med.2 <-
  brm(family = gaussian,
      y_model + m_model + set_rescor(FALSE),
       prior = c(prior(normal(0, 1), class = Intercept, resp = M),
                 prior(normal(0, 1), class = Intercept, resp = Y),
                prior(normal(0, 2), class = b, coef = X, resp = M),
                prior(normal(0, 2), class = b, coef = M, resp = Y),
                prior(normal(0, 2), class = b, coef = X, resp = Y),
                prior(exponential(1), class = sigma, resp = M),
                prior(exponential(1), class = sigma, resp = Y)),
      data = Data,
      cores = 4,
      file = "med.2")
```

---
```{r}
summary(med.2)
```

---

```{r}
med.2 %>% 
  spread_draws(b_Y_M, b_M_X) %>% 
  mutate(indirect = b_Y_M * b_M_X) %>% 
  median_qi(indirect)
```


```{r}
med.1 %>% 
  spread_draws(b_Y_M, b_M_X) %>% 
  mutate(indirect = b_Y_M * b_M_X) %>% 
  median_qi(indirect)
```

---
# Multiple Predictors, mediators and outcomes

```{r}
n <- 1e3
set.seed(4.5)
mult.X <-
  tibble(X1 = rnorm(n, mean = 0, sd = 1),
         X2 = rnorm(n, mean = 0, sd = 1),
         X3 = rnorm(n, mean = 0, sd = 1)) %>% 
  mutate(med = rnorm(n, mean = 0 + X1 * -1 + X2 * 0 + X3 * 1, sd = 1),
         dv  = rnorm(n, mean = 0 + X1 * 0 + X2 * .5 + X3 * 1 + M * .5, sd = 1))

```


```{r, echo = FALSE}
MultX_coords <-
  tibble(name = c("X1","X2","X3",  "M", "Y"),
         x    = c(1,1,1, 2, 3),
         y    = c(2,1,3, 1, 2))

X1 <-
  dagify(M ~ X1 + X2 + X3,
       Y ~ X1 + X2 + X3 + M,
       coords = MultX_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "black", alpha = 1/4, size = 10) +
  geom_dag_text(color = "black") +
  geom_dag_edges(edge_color = "black") +
  scale_x_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  theme_bw() +
  theme(panel.grid = element_blank())+
  ggtitle("direct + indirect effect")
X1
```

---

```{r}
med.3 <-
  brm(family = gaussian,
      bf(dv ~ 1 + X1 + X2 + X3 + med) + 
        bf(med ~ 1 + X1 + X2 + X3) + 
        set_rescor(FALSE),
      data = mult.X, 
      file = "med.3", 
      cores = 4)
```

---
```{r}
summary(med.3)
```


---
## Multiple outcomes

```{r}
n <- 1e3

set.seed(4.5)
Ys <-
  tibble(X  = rnorm(n, mean = 0, sd = 1)) %>% 
  mutate(M = rnorm(n, mean = 0 + X * .5, sd = 1)) %>% 
  mutate(Y1 = rnorm(n, mean = 0 + X * -1 + M * 0,  sd = 1),
         Y2 = rnorm(n, mean = 0 + X * 0  + M * .5, sd = 1),
         Y3 = rnorm(n, mean = 0 + X * 1  + M * 1,  sd = 1))

```


```{r, echo = FALSE}
MultY_coords <-
  tibble(name = c("X","M","Y1",  "Y2", "Y3"),
         x    = c(1,2,3, 3, 3),
         y    = c(2,2.25,1, 2, 3))

Y1 <-
  dagify(M ~ X,
       Y1 ~ X +  M,
       Y2 ~ X +  M,
       Y3 ~ X +  M,
       coords = MultY_coords) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "black", alpha = 1/4, size = 10) +
  geom_dag_text(color = "black") +
  geom_dag_edges(edge_color = "black") +
  scale_x_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  theme_bw() +
  theme(panel.grid = element_blank())+
  ggtitle("direct + indirect effect")
Y1

```

---
```{r}
mult.Ys <-
  brm(family = gaussian,
      bf(Y1 ~ 1 + X + M) + 
        bf(Y2 ~ 1 + X + M) + 
        bf(Y3 ~ 1 + X + M) + 
        bf(M ~ 1 + X) + 
        set_rescor(FALSE),
      data = Ys, 
      cores = 4,
      fit = "med.4")
```

---
.small[
```{r}
summary(mult.Ys)
```
]

---
## Multiple Mediators

parallel vs serial

```{r, echo = FALSE}
parallel <-
  tibble(name = c("X","M1","M2",  "Y"),
         x    = c(1,2,2,3),
         y    = c(2,2.25,1.75, 2))


 X2<- dagify(M1 ~ X,
       M2 ~ X ,
       Y ~ X +  M1,
       Y ~ X +  M2,
       coords = parallel) %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "black", alpha = 1/4, size = 10) +
  geom_dag_text(color = "black") +
  geom_dag_edges(edge_color = "black") +
  scale_x_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  theme_bw() +
  theme(panel.grid = element_blank())+
  ggtitle("parallel")
 
 
 serial <-
  tibble(name = c("X","M1","M2",  "Y"),
         x    = c(1,1.5,2.5,3),
         y    = c(2,2.25,2.25, 2))
 
  X3<- dagify(M1 ~ X,
       M2 ~ M1 ,
        M2 ~ X ,
       Y ~ M1,
       Y ~ X +  M2,
       coords = serial)  %>%
  
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(color = "black", alpha = 1/4, size = 10) +
  geom_dag_text(color = "black") +
  geom_dag_edges(edge_color = "black") +
  scale_x_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  scale_y_continuous(NULL, breaks = NULL, expand = c(.1, .1)) +
  theme_bw() +
  theme(panel.grid = element_blank())+
  ggtitle("serial")

X2 | X3
```
Note we can compute individual and total indirect effects

---
## parallel
```{r, eval = FALSE}
m1_model <- bf(M1 ~ 1 + X)
m2_model <- bf(M2  ~ 1 + X)
y_model  <- bf(Y ~ 1 + X + M1 + M2)

par <-
  brm(family = gaussian,
      y_model + m1_model + m2_model + set_rescor(FALSE),
      data = d)
```
3 indirect effects can be calculated
---
## serial
```{r, eval = FALSE}
ser <-
  brm(family = gaussian,
        bf(M1 ~ 1 + X) + 
        bf(M2 ~ 1 + X + M1) + 
        bf(Y ~ 1 + X + M1 + M2) + 
        set_rescor(FALSE),
        data=d)

```

4 indirect effects can be calculated

---
## moderated mediation? 
```{r, eval = FALSE}
y_model <- bf(Y ~ 1 + X + M)
m_model <- bf(M ~ 1 + X*moderator)

med.5 <-
  brm(family = gaussian,
      y_model + m_model + set_rescor(FALSE),
      data = Data,
      cores = 4,
      file = "med.5")
```

---
# Multivariate models



---
# Distributional Models

